The hydra and the rooster

### 2020-08-16 latex/chapter-introduction.tex
new exercise

In our development, we adapt the definitions and prove many theorems which
we found in the following articles. 
\begin{itemize}
\item ``Accessible independence results for Peano arithmetic''  by Laurie Kirby and Jeff Paris~\cite{KP82}
\item ''Rapidly growing Ramsey Functions'' by Jussi Ketonen and Robert Solovay~\cite{KS81}
\item ``The Termite and the Tower'', by Will Sladek~\cite{Sladek07thetermite}
\item Chapter V of ``Proof Theory'' by Kurt Schütte~\cite{Schutte}
\end{itemize}

@@

\subsubsection*{State of the development}
The \coq{} scripts herein are in constant development since our contribution~\cite{CantorContrib} on ordinal notations for the ordinals $\epsilon_0$ and $\Gamma_0$.
We added new material : axiomatic definition of countable ordinals after Schütte~\cite{Schutte}, combinatorial aspects of $\epsilon_0$, after Ketonen and Solovay~\cite{KS81} and Kirby and Paris~\cite{KP82}, recent \coq{} technology: type classes, equations, etc.

>>>>

In our development, we adapt the definitions and prove many theorems which
we found in the following articles. 
\begin{itemize}
\item ``Accessible independence results for Peano arithmetic''  by Laurie Kirby and Jeff Paris~\cite{KP82}
\item ''Rapidly growing Ramsey Functions'' by Jussi Ketonen and Robert Solovay~\cite{KS81}
\item ``The Termite and the Tower'', by Will Sladek~\cite{Sladek07thetermite}
\item Chapter V of ``Proof Theory'' by Kurt Schütte~\cite{schutte}
\end{itemize}

@@

\subsubsection*{State of the development}
The \coq{} scripts herein are in constant development since our contribution~\cite{CantorContrib} on ordinal notations for the ordinals $\epsilon_0$ and $\Gamma_0$.
We added new material : axiomatic definition of countable ordinals after Schütte~\cite{schutte}, combinatorial aspects of $\epsilon_0$, after Ketonen and Solovay~\cite{KS81} and Kirby and Paris~\cite{KP82}, recent \coq{} technology: type classes, equations, etc.

### 2020-08-16 latex/chapter-large-sets.tex
new exercise

But we do not know how to fill the dots yet \dots{}   In the next section, we will 
use \coq{} to reason  about the \emph{specification} of \texttt{L},
prove properties of any function which satisfies this specification.
In Sect.~\ref{sect:L-equaions}, we use the \texttt{coq-equations} plug-in
to define a function \texttt{L\_}, and prove its correctness w.r.t. its specification.

@@

\begin{itemize}
\item If $\alpha=0$, then $\H_\alpha (k)= k$ for any natural number $k$.
\item If $\alpha=\textrm{succ}(\beta)$, then 
$H_\alpha(k)=H_\beta(k+1)$ for any $k \in \mathbb{N}$
\item If $\alpha$ is a limit ordinal, then 
$\H_\alpha(k) = H_{(\canonseq{\alpha}{k+1})}(k)$ for any $k\in \mathbb{N}$.
\end{itemize}

>>>>

But we do not know how to fill the dots yet \dots{}   In the next section, we will 
use \coq{} to reason  about the \emph{specification} of \texttt{L},
prove properties of any function which satisfies this specification.
In Sect.~\ref{sect:L-equations}, we use the \texttt{coq-equations} plug-in
to define a function \texttt{L\_}, and prove its correctness w.r.t. its specification.

@@

\begin{itemize}
\item If $\alpha=0$, then $H_\alpha (k)= k$ for any natural number $k$.
\item If $\alpha=\textrm{succ}(\beta)$, then 
$H_\alpha(k)=H_\beta(k+1)$ for any $k \in \mathbb{N}$
\item If $\alpha$ is a limit ordinal, then 
$H_\alpha(k) = H_{(\canonseq{\alpha}{k+1})}(k)$ for any $k\in \mathbb{N}$.
\end{itemize}

### 2020-08-16 latex/chapter-termination.tex
new exercise

\index{Coq!Techniques!Coercions}
\index{Functions!Coercions@Coercions (from nat to ordinal types)}
\begin{remark}
Please refer to the remark~\pageref{waening:coercions} about the use of coercions.
% The use of coercions like \texttt{fin} allow us to be close to the mathematical tradition where natural numbers are ordinals too.
% Nevertheless, it may happen that a goal like \texttt{3 < 5} could be 
% interpreted as \texttt{(lt (fin 3) (fin 5))},  depending on the current notation scope.  
% When this misinterpretation happens, tactics like \texttt{auto with arith}, \texttt{lia} do not work!
% Thus, it is useful to write \texttt{(3 < 5)\%nat}  an inequality between two natural numbers. 
\end{remark}

@@

The type \texttt{T1} is not a correct ordinal notation, since it contains terms that are not in Cantor normal form. Moreover,
in Sections~\ref{sect:equations}, \ref{sect:hardy},
and \ref{sect:wainer}, we needed to define total recursive functions on well-formed ordinal terms less  than $\epsilon_0$, using the \texttt{Equations} plug-in~\cite{sozeau:hal-01671777}.
 In order to define a type whose inhabitants represent just ordinals, we build a type gathering a term of type \texttt{T1} and a proof that this term is in normal form.

@@

 Likewise, we define an addition in type \texttt{E0}, using the function{T1.plus}
and its properties.

@@

>>>>

% \index{Coq!Techniques!Coercions}
% \index{Functions!Coercions@Coercions (from nat to ordinal types)}
% \begin{remark}
% Please refer to the remark~\pageref{warning:coercions} about the use of coercions.
% % The use of coercions like \texttt{fin} allow us to be close to the mathematical tradition where natural numbers are ordinals too.
% % Nevertheless, it may happen that a goal like \texttt{3 < 5} could be 
% % interpreted as \texttt{(lt (fin 3) (fin 5))},  depending on the current notation scope.  
% % When this misinterpretation happens, tactics like \texttt{auto with arith}, \texttt{lia} do not work!
% % Thus, it is useful to write \texttt{(3 < 5)\%nat}  an inequality between two natural numbers. 
% \end{remark}

@@

The type \texttt{T1} is not a correct ordinal notation, since it contains terms that are not in Cantor normal form. Moreover,
in Sections~\ref{sect:L-equations}, \ref{sect:hardy},
and \ref{sect:wainer}, we needed to define total recursive functions on well-formed ordinal terms less  than $\epsilon_0$, using the \texttt{Equations} plug-in~\cite{sozeau:hal-01671777}.
 In order to define a type whose inhabitants represent just ordinals, we build a type gathering a term of type \texttt{T1} and a proof that this term is in normal form.

@@

@@

\subsubsection{Arithmetic on type \texttt{E0}}

 We define an addition in type \texttt{E0}, using the function{T1.plus}
and its properties.

\index{Exercises}
 \begin{exercise}
Let $\alpha$ be an ordinal. We say that $\alpha$ is \emph{infinite} iff the segment $[0,\alpha)$ is an infinite set. 
\begin{enumerate}
\item Adapt this definition to the type \texttt{E0}.
\item Prove that \emph{being infinite} is decidable on type \texttt{E0}.
\item Prove that $\alpha$ is infinite if and only if for all finite ordinal $i$,
$i+\alpha=\alpha$.
\end{enumerate}
 \end{exercise}
